---
title: "Rent burden for children under 5"
description: |
  58.6% of children under 5 are rent burdened.
author:
  - name: Catalyst California
    url: catalystcalifornia.org
date: 2024-02-13
output:
  distill::distill_article:
    css: css_theme.css
    self_contained: false
---


```{r setup, include=FALSE}

# get LA County under5 ages 0-4 by race in rent-burdened households
# trying this different cut that last year's households with under5 0-4 that are rent-burdened
# Data Dictionary: https://www2.census.gov/programs-surveys/acs/tech_docs/pums/data_dict/PUMS_Data_Dictionary_2018-2022.pdf

library(tidyverse)
library(data.table)
library(readxl)
library(tidycensus)
library(srvyr)
library(highcharter)
library(stringr)

#SOURCE from the script that has: styling, packages, dbconnection, colors

source("W:\\RDA Team\\R\\credentials_source.R")
con <- connect_to_db("f5la_v2")

source("W:\\Project\\RDA Team\\First5LA\\First 5 LA Data Hub Year 5 2024\\R\\chart_prep\\f5la_chart_functions.R")

```

```{r data, include=FALSE}

#  Step 1: load data

# PUMS Data
root <- "W:/Data/Demographics/PUMS/"

# Load the people PUMS data
people <- fread(paste0(root, "CA_2018_2022/psam_p06.csv"), header = TRUE, data.table = FALSE,
             colClasses = list(character = c("PUMA10", "PUMA20", "ANC1P", "ANC2P", "HISP", "RAC1P", "RAC2P", "RAC3P", "RACAIAN", "RACPI", "RACNH")))


# Load the housing PUMS data
housing <- fread(paste0(root, "CA_2018_2022/psam_h06.csv"), header = TRUE, data.table = FALSE,
                         colClasses = list(character = c("PUMA10", "PUMA20")))


#  Step 2: filter LA County under5 and re-code race/ethnicity 

# Select LA County under 5
la_cnty_under5 <- people %>% filter(AGEP <= 4 & (grepl('037', PUMA20) | grepl('037', PUMA10)))

# Sum under5 to test accuracy
#N <- sum(la_cnty_under5$PWGTP) #536815 close enough to children under 5 5-year avg


# Recode race/eth
la_cnty_under5<-la_cnty_under5%>%
  mutate(latino=(ifelse(HISP %in% "01", "Not Latinx", "Latinx")))

la_cnty_under5<-la_cnty_under5%>%
  mutate(aian=(ifelse(RACAIAN %in% "0", "Not AIAN", "AIAN Alone or in Combination")))

la_cnty_under5<-la_cnty_under5%>%
  mutate(nhpi=(ifelse(RACPI %in% "1", "NHPI Alone or in Combination",
                      ifelse(RACNH %in% '1', 'NHPI Alone or in Combination', "Not NHPI"))))


###### Set up codes for SWANA ######
# First use ancestry codes to help identify estimated swana pop
## Create list of swana codes for PUMS
pums_swana_list<-list("Algerian","Arab","Assyrian","Bahraini","Berber","Chaldean","Egyptian","Emirati","Iranian","Iraqi","Israeli","Jordanian","Kurdish","Kuwaiti","Lebanese","Libyan","Middle Eastern","Moroccan","North African","Omani","Palestinian","Qatari","Saudi","Syriac","Syrian","Tunisian","Yazidi","Yemeni","Mideast","Saudi Arabian","Arabic","Other Arab","Libyan (2017 or later)","Kuwaiti (2017 or later)","Turkish","Sudanese","Afghan") # 2017 or later needed based on reviewing data dictionary and saw fields for Arabic and Other Arab

## import PUMS codes
pums_codes <- read_excel("W:/Data/Demographics/PUMS/CA_2018_2022/PUMS_Data_Dictionary_2018-2022_ANC1P.xlsx")%>%
  mutate_all(as.character) # create this excel document separate by opening PUMS Data Dictionary in excel and deleting everything but ancestry fields -- since ANC1P and ANC2P have same data values no need to do both
pums_codes <- pums_codes %>% dplyr::rename("ANC_Code" = "Code_1")

## filter PUMS codes for swana descriptions based on our swana_list
swana_codes<-pums_codes%>%filter(Description %in% pums_swana_list)

##swana - alone or in combination with another race or latino
la_cnty_under5$swana <- "Not SWANA"
la_cnty_under5$swana[la_cnty_under5$ANC1P%in% swana_codes$ANC_Code| la_cnty_under5$ANC2P%in% swana_codes$ANC_Code] <- "SWANA"
la_cnty_under5$swana <- as.factor(la_cnty_under5$swana)

# code other race groups
la_cnty_under5$race = as.factor(ifelse(la_cnty_under5$RAC1P == "1" & la_cnty_under5$latino =="Not Latinx", "White NL",
                                     ifelse(la_cnty_under5$RAC1P == "1" & la_cnty_under5$latino =="Latinx", "Latinx placeholder",
                                            ifelse(la_cnty_under5$RAC1P == "2" & la_cnty_under5$latino =="Not Latinx", "Black NL",
                                                   ifelse(la_cnty_under5$RAC1P== "3" | la_cnty_under5$RAC1P== "4"|la_cnty_under5$RAC1P== "5", "AIAN placeholder",
                                                          ifelse(la_cnty_under5$RAC1P == "6" & la_cnty_under5$latino =="Not Latinx", "Asian NL",
                                                                 ifelse(la_cnty_under5$RAC1P == "7", "NHPI placeholder",
                                                                        ifelse(la_cnty_under5$RAC1P == "8" & la_cnty_under5$latino =="Not Latinx", "Other NL", 
                                                                               ifelse(la_cnty_under5$RAC1P== "9" 
                                                                                      & la_cnty_under5$latino =="Not Latinx", "Two or More NL", "Latinx placeholder"))))))))) 


la_cnty_under5<-la_cnty_under5%>%
  mutate(bipoc=(ifelse(race == "White NL" & swana == "Not SWAWNA", "Not BIPOC", "BIPOC")))

#test <- la_cnty_under5 %>% select(SERIALNO, HISP, RAC1P, race, aian, latino, nhpi, swana, bipoc) 
#View(test)



# trying out Asian disagg

# First use ancestry codes to help identify estimated asian pop
## Create list of asian codes for PUMS
pums_asian_list<-list("Afghan","Bangladeshi","Bhutanese","Nepali","Asian Indian","Bengali","East Indian","Punjabi","Pakistani","Sri Lankan", "Burmese", "Cambodian", "Chinese", "Cantonese", "Mongolian", "Tibetan", "Hong King (2017 or later)", "Filipino", "Indonesian", "Japanese", "Okinawan", "Korean", "Latioan", "Hmong", "Malaysian", "Thai", "Taiwanese", "Vietnamese", "Asian", "Other Asian")

## PUMS codes imported earlier

## filter PUMS codes for asian descriptions based on our asian_list
asian_codes<-pums_codes%>%filter(Description %in% pums_asian_list)

##asian - alone or in combination with another race or latino
la_cnty_under5$asian <- "Not Asian"
la_cnty_under5$asian[la_cnty_under5$ANC1P%in% asian_codes$ANC_Code| la_cnty_under5$ANC2P%in% asian_codes$ANC_Code] <- "Asian"
la_cnty_under5$asian <- as.factor(la_cnty_under5$asian)

# code asian subgroups
la_cnty_under5$asian = as.factor(ifelse(la_cnty_under5$ANC1P == "600" | la_cnty_under5$ANC2P =="600", "Afghan",
                                ifelse(la_cnty_under5$ANC1P == "603" | la_cnty_under5$ANC2P =="603", "Bangladeshi",
                                ifelse(la_cnty_under5$ANC1P == "607" | la_cnty_under5$ANC2P =="607", "Bhutanese",
                                ifelse(la_cnty_under5$ANC1P == "609" | la_cnty_under5$ANC2P =="609", "Nepali",
                                ifelse(la_cnty_under5$ANC1P == "615" | la_cnty_under5$ANC2P =="615", "Indian",
                                ifelse(la_cnty_under5$ANC1P == "618" | la_cnty_under5$ANC2P =="618", "Bengali",
                                ifelse(la_cnty_under5$ANC1P == "620" | la_cnty_under5$ANC2P =="620", "East Indian",
                                ifelse(la_cnty_under5$ANC1P == "650" | la_cnty_under5$ANC2P =="650", "Punjabi",
                                ifelse(la_cnty_under5$ANC1P == "680" | la_cnty_under5$ANC2P =="680", "Pakistani",
                                ifelse(la_cnty_under5$ANC1P == "690" | la_cnty_under5$ANC2P =="690", "Sri Lankan",
                                ifelse(la_cnty_under5$ANC1P == "700" | la_cnty_under5$ANC2P =="700", "Burmese",
                                ifelse(la_cnty_under5$ANC1P == "703" | la_cnty_under5$ANC2P =="703", "Cambodian",
                                ifelse(la_cnty_under5$ANC1P == "706" | la_cnty_under5$ANC2P =="706", "Chinese",
                                ifelse(la_cnty_under5$ANC1P == "707" | la_cnty_under5$ANC2P =="707", "Cantonese",
                                ifelse(la_cnty_under5$ANC1P == "712" | la_cnty_under5$ANC2P =="712", "Mongolian",
                                ifelse(la_cnty_under5$ANC1P == "714" | la_cnty_under5$ANC2P =="714", "Tibetan",
                                ifelse(la_cnty_under5$ANC1P == "716" | la_cnty_under5$ANC2P =="716", "Hong Kong",
                                ifelse(la_cnty_under5$ANC1P == "720" | la_cnty_under5$ANC2P =="720", "Filipino",
                                ifelse(la_cnty_under5$ANC1P == "730" | la_cnty_under5$ANC2P =="730", "Indonesian",
                                ifelse(la_cnty_under5$ANC1P == "740" | la_cnty_under5$ANC2P =="740", "Japanese",
                                ifelse(la_cnty_under5$ANC1P == "748" | la_cnty_under5$ANC2P =="748", "Okinawan",
                                ifelse(la_cnty_under5$ANC1P == "750" | la_cnty_under5$ANC2P =="750", "Korean",
                                ifelse(la_cnty_under5$ANC1P == "765" | la_cnty_under5$ANC2P =="765", "Laotian",
                                ifelse(la_cnty_under5$ANC1P == "768" | la_cnty_under5$ANC2P =="768", "Hmong",
                                ifelse(la_cnty_under5$ANC1P == "770" | la_cnty_under5$ANC2P =="770", "Malaysian",
                                ifelse(la_cnty_under5$ANC1P == "776" | la_cnty_under5$ANC2P =="776", "Thai",
                                ifelse(la_cnty_under5$ANC1P == "782" | la_cnty_under5$ANC2P =="782", "Taiwanese",
                                ifelse(la_cnty_under5$ANC1P == "785" | la_cnty_under5$ANC2P =="785", "Vietnamese",
                                ifelse(la_cnty_under5$ANC1P == "795" | la_cnty_under5$ANC2P =="795", "Asian",
                                ifelse(la_cnty_under5$ANC1P == "799" | la_cnty_under5$ANC2P =="799", "Other Asian",
                                       "Not Asian"
                                       )))))))))))))))))))))))))))))))

#test <- la_cnty_under5 %>% select(SERIALNO, HISP, RAC1P, ANC1P, ANC2P, nhpi, asian) 
#View(test)


####  Step 3: filter LA County households eligible for rent-burden calculation and join to under5  #### 

## Select LA County renter households (in cost-burden universe)
la_cnty_renter_hhs <- housing %>% filter(TEN <= "3" & is.na(GRPIP) == FALSE & (grepl('037', PUMA20) | grepl('037', PUMA10)))

## Sum renter hhs to test accuracy
#N <- sum(la_cnty_renter_hhs$WGTP) #1,733,508 close enough to the 1,796,882 I see here: 
#https://data.census.gov/table/ACSDT5Y2021.B25070?q=B25070:%20Gross%20Rent%20as%20a%20Percentage%20of%20Household%20Income%20in%20the%20Past%2012%20Months&g=050XX00US06037&tid=ACSDT1Y2022.B25070

#join and filter
eligible_under5 <- la_cnty_under5 %>% left_join(la_cnty_renter_hhs, by = "SERIALNO") %>%
  filter(is.na(WGTP) == FALSE)

#sum(eligible_under5$PWGTP) = 298,784 seems about right if slightly more than half of under 5 are eligible (renters)


#### Step 4: set up surveys and calculate percentages of under5 in rent-burdened households by race/ethnicity

# survey design code

# Define weight variable and population base which will be used in the survey design set up
## You must use WGTP (if you are using psam_h06.csv and want housing units, like for Low Quality Housing) or PWGTP (if you want person units, like for Connected under5)
weight <- 'PWGTP' # using PWGTP b/c calculating percentage of under 5 in rent-burdened households (from psam_p06.csv)

repwlist = rep(paste0("PWGTP", 1:80))

# prep data and add in burdened indicator
under5 <- eligible_under5
under5$geoid <- "037"

under5<-under5%>%
  mutate(indicator=(ifelse(under5$GRPIP > 29, "burdened", "not burdened")))

# create survey design

under5_county <- under5 %>%               
  as_survey_rep(
    variables = c(geoid, indicator, race, latino, aian, nhpi, swana, bipoc, asian),   # dplyr::select grouping variables
    weights = weight,                       #  weight
    repweights = repwlist,                  # list of replicate weights
    combined_weights = TRUE,                # tells the function that replicate weights are included in the data
    mse = TRUE,                             # tells the function to calc mse
    type="other",                           # statistical method
    scale=4/80,                             # scaling set by ACS
    rscale=rep(1,80)                        # setting specific to ACS-scaling
  )

###### Latino ######
lat <- under5_county  %>%
  group_by(geoid,latino,indicator) %>%   # group by race cat
  summarise(
    num = survey_total(na.rm=T), # get the (survey weighted) count for the numerators
    rate = survey_mean()) %>%        # get the (survey weighted) proportion for the numerator
  left_join(under5_county %>%                                        # left join in the denominators
              group_by(geoid,latino) %>%                                     # group by geo
              summarise(pop = survey_total(na.rm=T))) %>%              # get the weighted total for overall geo
  mutate(rate=rate*100,
         rate_moe = rate_se*1.645*100,    # calculate the margin of error for the rate based on se
         rate_cv = ((rate_moe/1.645)/rate) * 100, # calculate cv for rate
         count_moe = num_se*1.645, # calculate moe for numerator count based on se
         count_cv = ((count_moe/1.645)/num) * 100)  # calculate cv for numerator count


###### NHPI ######
nhpi <- under5_county  %>%
  group_by(geoid,nhpi,indicator) %>%   # group by race cat
  summarise(
    num = survey_total(na.rm=T), # get the (survey weighted) count for the numerators
    rate = survey_mean()) %>%        # get the (survey weighted) proportion for the numerator
  left_join(under5_county %>%                                        # left join in the denominators
              group_by(geoid,nhpi) %>%                                     # group by geo
              summarise(pop = survey_total(na.rm=T))) %>%              # get the weighted total for overall geo
  mutate(rate=rate*100,
         rate_moe = rate_se*1.645*100,    # calculate the margin of error for the rate based on se
         rate_cv = ((rate_moe/1.645)/rate) * 100, # calculate cv for rate
         count_moe = num_se*1.645, # calculate moe for numerator count based on se
         count_cv = ((count_moe/1.645)/num) * 100)  # calculate cv for numerator count

###### AIAN ######
aian <- under5_county  %>%
  group_by(geoid,aian,indicator) %>%   # group by race cat
  summarise(
    num = survey_total(na.rm=T), # get the (survey weighted) count for the numerators
    rate = survey_mean()) %>%        # get the (survey weighted) proportion for the numerator
  left_join(under5_county %>%                                        # left join in the denominators
              group_by(geoid,aian) %>%                                     # group by geo
              summarise(pop = survey_total(na.rm=T))) %>%              # get the weighted total for overall geo
  mutate(rate=rate*100,
         rate_moe = rate_se*1.645*100,    # calculate the margin of error for the rate based on se
         rate_cv = ((rate_moe/1.645)/rate) * 100, # calculate cv for rate
         count_moe = num_se*1.645, # calculate moe for numerator count based on se
         count_cv = ((count_moe/1.645)/num) * 100)  # calculate cv for numerator count

###### SWANA ######
swana <- under5_county  %>%
  group_by(geoid,swana,indicator) %>%   # group by race cat
  summarise(
    num = survey_total(na.rm=T), # get the (survey weighted) count for the numerators
    rate = survey_mean()) %>%        # get the (survey weighted) proportion for the numerator
  left_join(under5_county %>%                                        # left join in the denominators
              group_by(geoid,swana) %>%                                     # group by geo
              summarise(pop = survey_total(na.rm=T))) %>%              # get the weighted total for overall geo
  mutate(rate=rate*100,
         rate_moe = rate_se*1.645*100,    # calculate the margin of error for the rate based on se
         rate_cv = ((rate_moe/1.645)/rate) * 100, # calculate cv for rate
         count_moe = num_se*1.645, # calculate moe for numerator count based on se
         count_cv = ((count_moe/1.645)/num) * 100)  # calculate cv for numerator count

###### RACE ######
race <- under5_county  %>%
  group_by(geoid,race,indicator) %>%   # group by race cat
  summarise(
    num = survey_total(na.rm=T), # get the (survey weighted) count for the numerators
    rate = survey_mean()) %>%        # get the (survey weighted) proportion for the numerator
  left_join(under5_county %>%                                        # left join in the denominators
              group_by(geoid,race) %>%                                     # group by geo
              summarise(pop = survey_total(na.rm=T))) %>%              # get the weighted total for overall geo
  mutate(rate=rate*100,
         rate_moe = rate_se*1.645*100,    # calculate the margin of error for the rate based on se
         rate_cv = ((rate_moe/1.645)/rate) * 100, # calculate cv for rate
         count_moe = num_se*1.645, # calculate moe for numerator count based on se
         count_cv = ((count_moe/1.645)/num) * 100)  # calculate cv for numerator count

###### BIPOC ######
bipoc <- under5_county  %>%
  group_by(geoid,bipoc,indicator) %>%   # group by race cat
  summarise(
    num = survey_total(na.rm=T), # get the (survey weighted) count for the numerators
    rate = survey_mean()) %>%        # get the (survey weighted) proportion for the numerator
  left_join(under5_county %>%                                        # left join in the denominators
              group_by(geoid,bipoc) %>%                                     # group by geo
              summarise(pop = survey_total(na.rm=T))) %>%              # get the weighted total for overall geo
  mutate(rate=rate*100,
         rate_moe = rate_se*1.645*100,    # calculate the margin of error for the rate based on se
         rate_cv = ((rate_moe/1.645)/rate) * 100, # calculate cv for rate
         count_moe = num_se*1.645, # calculate moe for numerator count based on se
         count_cv = ((count_moe/1.645)/num) * 100)  # calculate cv for numerator count

###### TOTAL ######
total <- under5_county  %>%
  group_by(geoid,indicator) %>%   # group by race cat
  summarise(
    num = survey_total(na.rm=T), # get the (survey weighted) count for the numerators
    rate = survey_mean()) %>%        # get the (survey weighted) proportion for the numerator
  left_join(under5_county %>%                                        # left join in the denominators
              group_by(geoid) %>%                                     # group by geo
              summarise(pop = survey_total(na.rm=T))) %>%              # get the weighted total for overall geo
  mutate(rate=rate*100,
         rate_moe = rate_se*1.645*100,    # calculate the margin of error for the rate based on se
         rate_cv = ((rate_moe/1.645)/rate) * 100, # calculate cv for rate
         count_moe = num_se*1.645, # calculate moe for numerator count based on se
         count_cv = ((count_moe/1.645)/num) * 100)  # calculate cv for numerator count

###### ASIAN DISAGG ######
asian <- under5_county  %>%
  group_by(geoid,asian,indicator) %>%   # group by race cat
  summarise(
    num = survey_total(na.rm=T), # get the (survey weighted) count for the numerators
    rate = survey_mean()) %>%        # get the (survey weighted) proportion for the numerator
  left_join(under5_county %>%                                        # left join in the denominators
              group_by(geoid,asian) %>%                                     # group by geo
              summarise(pop = survey_total(na.rm=T))) %>%              # get the weighted total for overall geo
  mutate(rate=rate*100,
         rate_moe = rate_se*1.645*100,    # calculate the margin of error for the rate based on se
         rate_cv = ((rate_moe/1.645)/rate) * 100, # calculate cv for rate
         count_moe = num_se*1.645, # calculate moe for numerator count based on se
         count_cv = ((count_moe/1.645)/num) * 100)  # calculate cv for numerator count


####  Step 5: format

# rename race name columns as subgroup
total$subgroup = "Total"
total <- total %>% select(geoid, subgroup, everything())

aian <- aian %>% rename(subgroup = aian)
bipoc <- bipoc %>% rename(subgroup = bipoc)
lat <- lat %>% rename(subgroup = latino)
nhpi <- nhpi %>% rename(subgroup = nhpi)
race <- race %>% rename(subgroup = race)
swana <- swana %>% rename(subgroup = swana)

# merge tables except for bipoc - need total
d_long <- rbind(total, aian, lat, race, nhpi, swana) %>%
  filter(indicator == "burdened" & 
           subgroup != "Not AIAN" &
           subgroup != "Not Latinx" &
           subgroup != "Not NHPI" &
           subgroup != "Not SWANA" &
           subgroup != "AIAN placeholder" &
           subgroup != "NHPI placeholder" &
           subgroup != "Latinx placeholder")
           
d_long <- as.data.frame(d_long)
d_long$geoid <- "06037"
d_long <- d_long %>% select(geoid, everything())


```


```{r chart, echo = FALSE}

#chart

#format for chart
d_chart <- d_long %>%
  arrange(desc(rate)) %>%
  mutate(across('subgroup', str_replace, ' NL', '')) %>%
  mutate(across('subgroup', str_replace, ' Alone or in Combination', '')) %>%
  rename(population = num) %>%
  mutate(group = "Percent of children under 5 in rent-burdened households")

hc <- d_chart %>%
  hchart('bar', hcaes(x = subgroup, y = rate, group = group)) %>%
  hc_yAxis(title = list(text = "")) %>%
  hc_xAxis(title = list(text = "")) %>%
  hc_tooltip(pointFormat = '{point.y:.1f}%') %>%
  hc_title(
    text = "SWANA, Black, and Latinx children under 5 are more likely to be rent burdened"
  ) %>%
   hc_exporting(
    enabled = TRUE, sourceWidth=900, sourceHeight=600,
    chartOptions=list(plotOptions=list(series=list(dataLabels=list(enabled=TRUE, format='{point.y:,.1f}%')))),
    filename = "plot"
  ) %>%
  hc_subtitle(text = "Percent of children under 5 in rent-burdened households by race") %>%
  hc_caption(text = "Source: ACS 2018-2022 5-year estimates <br> AIAN = American Indian or Alaska Native. NHPI = Native Hawaiian or Pacific Islander. SWANA = Southwest Asian or North African. In this chart White, Asian, Black, Other and Two or More are non-Hispanic subgroups; AIAN, NHPI, SWANA, and Two or More are alone or in combination with other races.") %>%
  hc_legend(enabled = F)
#hc
```


```{r chart 1, echo=FALSE,out.width="100%", out.height="100%"}

#TITLE 
title_text <- paste0("SWANA, Black, and Latinx children under 5 are more likely to be rent burdened")

#SUBTITLE
subtitle_text <- paste0("Percent of children age 0-5 experiencing rent burden in Los Angeles county")

#CAPTION
caption_text<- paste0(acsnote,"<br>",racenote) #can always customize or add additional text to this as needed

#TOOLTIP
#modify tooltips as needed 

tooltip_text <- paste0("<b>{point.rate:.1f}%</b> of <b>{point.subgroup}</b> people experience rent burden, or a total of <b>{point.pop:.1f}</b> people")


#STEP 4 RUN the hchart FUNCTION 

# source("W:\\Project\\RDA Team\\First5LA\\First 5 LA Data Hub Year 5 2024\\Github\\f5la_chart_functions.R")

graph <- fx_single(
    df = d_chart,
    x = 'subgroup',
    y = 'rate',
    chart_title = title_text,
    chart_subtitle = subtitle_text,
    chart_tooltip = tooltip_text,
    chart_caption = caption_text
)

graph


```


Note: Aggregated data above conceal variations with disaggregated subgroups. 100% of children under 5 of Afghan descent are rent burdened, and Hmong, Pakistani, Indonesian, and Bangladeshi children under 5 are all more likely to be rent burdened than LA County children overall. See chart below for details.

```{r chart2, echo = FALSE}

# # trying out Asian disagg

#chart

#format for chart
d_chart_asian <- asian %>% rename(subgroup = asian) %>%
  filter(indicator =="burdened" & pop >= 100 & rate_cv <= 40 & subgroup != "Asian" & subgroup != "Not Asian") %>%
  arrange(desc(rate)) %>%
  rename(population = num) %>%
  mutate(group = "Percent of children under 5 in rent-burdened households")

hc_asian <- d_chart_asian %>%
  hchart('column', hcaes(x = subgroup, y = rate, group = group)) %>%
  hc_tooltip(pointFormat = '{point.y:.1f}%') %>%
  hc_yAxis(title = list(text = ""), max = 100) %>%
  #hc_yAxis(title = list(text = "")) %>%
  hc_xAxis(title = list(text = "")) %>%
  hc_title(
    text = "Afghan children under 5 are most likely to be rent burdened"
  ) %>%
  hc_subtitle(text = "Percent of children under 5 in rent-burdened households by Asian ancestry") %>%
   hc_exporting(
    enabled = TRUE, sourceWidth=900, sourceHeight=600,
    chartOptions=list(plotOptions=list(series=list(dataLabels=list(enabled=TRUE, format='{point.y:,.1f}%')))),
    filename = "plot"
  ) %>%

  hc_caption(text = "Source: ACS 2018-2022 5-year estimates")%>%
  hc_legend(enabled = F)

hc_asian
```


## Why it matters?
First 5 LA's first strategic plan goal is that children prenatal to age 5 and their families have their basic needs met, which includes secure housing for essential psychological safety. This goal has an objective to reduce tot annual percent of households with children prenatal through age 5 who experience housing insecurity. Rent burden is not the sole cause of housing insecurity, but it is a key contributor.

## Learn more
See more in the [First 5 LA 2024-2029 Strategic Plan](https://www.first5la.org/wp-content/uploads/2023/11/First-5-LA-Strategic-Plan-2024-2029.pdf) and see related indicators of basic needs in the [Data Narrative](https://apca-dev.dreamhosters.com/).

<script>
  widg = document.getElementsByClassName('html-widget'); //find widgets
  for (let i = 0; i < widg.length; i++) { //loop through
    document.getElementsByClassName('highchart')[i].style.width = '100%'; //adjust width
    document.getElementsByClassName('highchart')[i].style.height = '100%'; //adjust height
  }
</script>


```{r, include = FALSE}
#Close postgres connection
dbDisconnect(con) 
```
