---
title: "Region 5 Indicators"
description: "Data related to children age 0 to 5 that align with Best Start needs and the new First 5 LA strategic plan within Region 5. Region 5 contains the Lancaster and Palmdale geographies."
author:
  - name: Catalyst California
    url: catalystcalifornia.org
date: 2024-06-05
categories:
  - Region 5
  - Indicators
output:
  distill::distill_article:
    toc: true
    toc_float: true
    css: css_theme.css
    self_contained: false
preview: images/reg5.PNG
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(highcharter)
options(scipen = 999)

# create connection for rda database and get data
source("W:\\RDA Team\\R\\credentials_source.R")
# f5la_con <- connect_to_db("f5la_v2")
# race_under5 <- dbGetQuery(f5la_con, "select * from data.f5la_2024_bs_raceethnicity_pums")
# poverty_under5 <- dbGetQuery(f5la_con, "select * from data.f5la_2024_bs_poverty_hh_pums")
# rent_burden_under5 <- dbGetQuery(f5la_con, "select * from data.f5la_2024_bs_rent_burden_hh_pums")

# dbDisconnect(f5la_con)

source("W:\\Project\\RDA Team\\First5LA\\First 5 LA Data Hub Year 5 2024\\R\\chart_prep\\f5la_chart_functions.R")

# JZ update 6/6/24: Outsource all the functions

source("W:\\Project\\RDA Team\\First5LA\\First 5 LA Data Hub Year 5 2024\\R\\analysis\\region_5_indicators_hh.R")

```




# Lancaster

## Race-Ethnicity

```{r race1}

best_start <- "Lancaster"
make_race_chart(race_under5, best_start)

```


## Poverty

```{r poverty1, out.width="100%", out.height="100%", results = 'asis'}

best_start <- "Lancaster"
make_poverty_chart(poverty_under5, best_start)

```


## Rent Burden

```{r rb1}

best_start <- "Lancaster"
make_rent_burden_chart(rent_burden_under5, best_start)

```



# Palmdale

## Race-Ethnicity

```{r race2}

best_start <- "Palmdale"
make_race_chart(race_under5, best_start)

```

## Poverty

```{r poverty2, out.width="100%", out.height="100%", results = 'asis'}

best_start <- "Palmdale"
make_poverty_chart(poverty_under5, best_start)

```

## Rent Burden

```{r rb2 chart}

best_start <- "Palmdale"
make_rent_burden_chart(rent_burden_under5, best_start)

```


# Why it matters
Understanding the racial-ethnic breakdown of an area is important to understanding its cultures and needs. Poverty and rent burden are indicators of economic need First 5 LA is interested in as it works to ensure all children prenatal to age 5 and their families have their basic needs met.

# Learn more
Learn more by reading the [First 5 LA 2024-2029 Strategic Plan](/https://www.first5la.org/wp-content/uploads/2023/11/First-5-LA-Strategic-Plan-2024-2029.pdf).

