---
title: "Central Long Beach children under 5 by race"
description: |
  
author:
  - name: Catalyst California
    url: catalystcalifornia.org
date: 2024-02-08
output:
  distill::distill_article:
    css: css_theme.css
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(data.table)
library(readxl)
library(tidycensus)
library(srvyr)
library(highcharter)
library(stringr)
library(sf)
library(tigris)
options(scipen = 999)

# create connection for rda database
source("W:\\RDA Team\\R\\credentials_source.R")

#get 2010 vintage PUMA xwalk
f5la_con <- connect_to_db("f5la_v2")
xwalk10 <- st_read(f5la_con, query = "select * from data.bsc_puma_crosswalk_2022")
dbDisconnect(f5la_con)

bsc_puma_xwalk10 <- as.data.frame(xwalk10) %>% mutate(PUMA10 = substr(geoid10, 3, 7)) %>%
  rename(pct_area_intersect10 = pct_area_intersect,
         pct_area_intersect_bsc10 = pct_area_intersect_bsc) %>%
select(best_start, PUMA10, pct_area_intersect10, pct_area_intersect_bsc10) 
  

#get best start geographies, convert to 3310, and calculate_area
best_starts <- st_read("W:/Project/RDA Team/First5LA/Shapes/f5la_best_start_geographies/f5la_best_start_geographies.shp")
best_starts <- st_transform(best_starts, 3310)
best_starts$bs_area_3310 <- st_area(best_starts)

#get pumas and filter for la and convert to 3310
pumas <- pumas(state = "CA", year = 2022) 
la_pumas <- pumas %>% filter(grepl('037', GEOID20))
la_pumas <- st_transform(la_pumas, 3310)
la_pumas$puma20_area_3310 <- st_area(la_pumas)

#intersect, calculate area, and filter based on area (5% or more of PUMA, 
#which ends up as 10% or more of the best start)
xwalk <- st_intersection(best_starts, la_pumas)
xwalk$intersect_area_3310 <- st_area(xwalk)
xwalk <- xwalk %>% mutate(pct_int_of_bs = intersect_area_3310 / bs_area_3310,
                          pct_int_of_puma20 = intersect_area_3310 / puma20_area_3310) %>%
  transform(pct_int_of_bs = as.numeric(pct_int_of_bs),
            pct_int_of_puma20 = as.numeric(pct_int_of_puma20))

#final
bsc_puma_xwalk20 <- as.data.frame(xwalk) %>% filter(pct_int_of_puma20 >= .05) %>%
  select(BEST_START, GEO_ID, PUMACE20, GEOID20, NAMELSAD20, pct_int_of_bs, pct_int_of_puma20) %>%
  rename(PUMA20 = PUMACE20)

```


```{r data}

# PUMS Data
root <- "W:/Data/Demographics/PUMS/"

## Load the people PUMS data
people <- fread(paste0(root, "CA_2018_2022/psam_p06.csv"), header = TRUE, data.table = FALSE,
             colClasses = list(character = c("PUMA10", "PUMA20", "ANC1P", "ANC2P", "HISP", "RAC1P", "RAC2P", "RAC3P", "RACAIAN", "RACPI", "RACNH")))

## Select LA County
la <- people %>% filter(grepl('037', PUMA20) | grepl('037', PUMA10) ) 

la <- la %>% left_join(bsc_puma_xwalk20, by = "PUMA20")
la <- la %>% left_join(bsc_puma_xwalk10, by = "PUMA10")

#create one best start field combining both vintages
la <- la %>% mutate(best_start_final = ifelse(is.na(la$best_start) & is.na(la$BEST_START), NA, 
                                                  ifelse(is.na(best_start), la$BEST_START,la$best_start)))


## Recode race/eth
la<-la%>%
  mutate(latino=(ifelse(HISP %in% "01", "Not Latinx", "Latinx")))

la<-la%>%
  mutate(aian=(ifelse(RACAIAN %in% "0", "Not AIAN", "AIAN Alone or in Combination")))

la<-la%>%
  mutate(nhpi=(ifelse(RACPI %in% "1", "NHPI Alone or in Combination",
                      ifelse(RACNH %in% '1', 'NHPI Alone or in Combination', "Not NHPI"))))

###### Set up codes for SWANA ######
# First use ancestry codes to help identify estimated swana pop
## Create list of swana codes for PUMS
pums_swana_list<-list("Algerian","Arab","Assyrian","Bahraini","Berber","Chaldean","Egyptian","Emirati","Iranian","Iraqi","Israeli","Jordanian","Kurdish","Kuwaiti","Lebanese","Libyan","Middle Eastern","Moroccan","North African","Omani","Palestinian","Qatari","Saudi","Syriac","Syrian","Tunisian","Yazidi","Yemeni","Mideast","Saudi Arabian","Arabic","Other Arab","Libyan (2017 or later)","Kuwaiti (2017 or later)","Turkish","Sudanese","Afghan") # 2017 or later needed based on reviewing data dictionary and saw fields for Arabic and Other Arab

## import PUMS codes
pums_codes <- read_excel("W:/Data/Demographics/PUMS/CA_2018_2022/PUMS_Data_Dictionary_2018-2022_ANC1P.xlsx")%>%
  mutate_all(as.character) # create this excel document separate by opening PUMS Data Dictionary in excel and deleting everything but ancestry fields -- since ANC1P and ANC2P have same data values no need to do both
pums_codes <- pums_codes %>% dplyr::rename("ANC_Code" = "Code_1")

## filter PUMS codes for swana descriptions based on our swana_list
swana_codes<-pums_codes%>%filter(Description %in% pums_swana_list)

##swana - alone or in combination with another race or latino
la$swana <- "Not SWANA"
la$swana[la$ANC1P%in% swana_codes$ANC_Code| la$ANC2P%in% swana_codes$ANC_Code] <- "SWANA"
la$swana <- as.factor(la$swana)

# code other race groups
la$race <- NA #declare variable to avoid unequal rows error
la$race = as.factor(ifelse(la$RAC1P == "1" & la$latino =="Not Latinx", "White NL",
                            ifelse(la$RAC1P == "1" & la$latino =="Latinx", "Latinx placeholder",
                            ifelse(la$RAC1P == "2" & la$latino =="Not Latinx", "Black NL",
                            ifelse(la$RAC1P == "3" | la$RAC1P== "4"|la$RAC1P== "5", "AIAN placeholder",
                            ifelse(la$RAC1P == "6" & la$latino =="Not Latinx", "Asian NL",
                            ifelse(la$RAC1P == "7" & la$latino =="Not Latinx", "NHPI placeholder",
                            ifelse(la$RAC1P == "8" & la$latino =="Not Latinx", "Other NL", 
                            ifelse(la$RAC1P == "9" & la$latino =="Not Latinx", "Two or More NL", "Latinx placeholder"))))))))) 

```


```{r}

# survey design code

# Define weight variable and population base which will be used in the survey design set up
## You must use WGTP (if you are using psam_h06.csv and want housing units, like for Low Quality Housing) or PWGTP (if you want person units)
weight <- 'PWGTP'

repwlist = rep(paste0("PWGTP", 1:80))
youth <- la
youth$geoid <- la$best_start_final

youth<-youth%>%
  mutate(indicator=(ifelse(youth$AGEP >= 5, "not under 5", "under 5")))

# create survey design
youth_bs <- youth %>%               
  as_survey_rep(
    variables = c(geoid, indicator, race, latino, aian, nhpi, swana),   # dplyr::select grouping variables
    weights = weight,                       #  weight
    repweights = repwlist,                  # list of replicate weights
    combined_weights = TRUE,                # tells the function that replicate weights are included in the data
    mse = TRUE,                             # tells the function to calc mse
    type="other",                           # statistical method
    scale=4/80,                             # scaling set by ACS
    rscale=rep(1,80)                        # setting specific to ACS-scaling
  )

###### Latino ######
lat <- youth_bs  %>%
  group_by(geoid,latino,indicator) %>%   # group by race cat
  summarise(
    num = survey_total(na.rm=T), # get the (survey weighted) count for the numerators
    rate = survey_mean()) %>%        # get the (survey weighted) proportion for the numerator
  left_join(youth_bs %>%                                        # left join in the denominators
              group_by(geoid,latino) %>%                                     # group by geo
              summarise(pop = survey_total(na.rm=T))) %>%              # get the weighted total for overall geo
  mutate(rate=rate*100,
         rate_moe = rate_se*1.645*100,    # calculate the margin of error for the rate based on se
         rate_cv = ((rate_moe/1.645)/rate) * 100, # calculate cv for rate
         count_moe = num_se*1.645, # calculate moe for numerator count based on se
         count_cv = ((count_moe/1.645)/num) * 100)  # calculate cv for numerator count

###### NHPI ######
nhpi <- youth_bs  %>%
  group_by(geoid,nhpi,indicator) %>%   # group by race cat
  summarise(
    num = survey_total(na.rm=T), # get the (survey weighted) count for the numerators
    rate = survey_mean()) %>%        # get the (survey weighted) proportion for the numerator
  left_join(youth_bs %>%                                        # left join in the denominators
              group_by(geoid,nhpi) %>%                                     # group by geo
              summarise(pop = survey_total(na.rm=T))) %>%              # get the weighted total for overall geo
  mutate(rate=rate*100,
         rate_moe = rate_se*1.645*100,    # calculate the margin of error for the rate based on se
         rate_cv = ((rate_moe/1.645)/rate) * 100, # calculate cv for rate
         count_moe = num_se*1.645, # calculate moe for numerator count based on se
         count_cv = ((count_moe/1.645)/num) * 100)  # calculate cv for numerator count

###### AIAN ######
aian <- youth_bs  %>%
  group_by(geoid,aian,indicator) %>%   # group by race cat
  summarise(
    num = survey_total(na.rm=T), # get the (survey weighted) count for the numerators
    rate = survey_mean()) %>%        # get the (survey weighted) proportion for the numerator
  left_join(youth_bs %>%                                        # left join in the denominators
              group_by(geoid,aian) %>%                                     # group by geo
              summarise(pop = survey_total(na.rm=T))) %>%              # get the weighted total for overall geo
  mutate(rate=rate*100,
         rate_moe = rate_se*1.645*100,    # calculate the margin of error for the rate based on se
         rate_cv = ((rate_moe/1.645)/rate) * 100, # calculate cv for rate
         count_moe = num_se*1.645, # calculate moe for numerator count based on se
         count_cv = ((count_moe/1.645)/num) * 100)  # calculate cv for numerator count

###### SWANA ######
swana <- youth_bs  %>%
  group_by(geoid,swana,indicator) %>%   # group by race cat
  summarise(
    num = survey_total(na.rm=T), # get the (survey weighted) count for the numerators
    rate = survey_mean()) %>%        # get the (survey weighted) proportion for the numerator
  left_join(youth_bs %>%                                        # left join in the denominators
              group_by(geoid,swana) %>%                                     # group by geo
              summarise(pop = survey_total(na.rm=T))) %>%              # get the weighted total for overall geo
  mutate(rate=rate*100,
         rate_moe = rate_se*1.645*100,    # calculate the margin of error for the rate based on se
         rate_cv = ((rate_moe/1.645)/rate) * 100, # calculate cv for rate
         count_moe = num_se*1.645, # calculate moe for numerator count based on se
         count_cv = ((count_moe/1.645)/num) * 100)  # calculate cv for numerator count

###### RACE ######
race <- youth_bs  %>%
  group_by(geoid,race,indicator) %>%   # group by race cat
  summarise(
    num = survey_total(na.rm=T), # get the (survey weighted) count for the numerators
    rate = survey_mean()) %>%        # get the (survey weighted) proportion for the numerator
  left_join(youth_bs %>%                                        # left join in the denominators
              group_by(geoid,race) %>%                                     # group by geo
              summarise(pop = survey_total(na.rm=T))) %>%              # get the weighted total for overall geo
  mutate(rate=rate*100,
         rate_moe = rate_se*1.645*100,    # calculate the margin of error for the rate based on se
         rate_cv = ((rate_moe/1.645)/rate) * 100, # calculate cv for rate
         count_moe = num_se*1.645, # calculate moe for numerator count based on se
         count_cv = ((count_moe/1.645)/num) * 100)  # calculate cv for numerator count

###### TOTAL ######
total <- youth_bs  %>%
  group_by(geoid,indicator) %>%   # group by race cat
  summarise(
    num = survey_total(na.rm=T), # get the (survey weighted) count for the numerators
    rate = survey_mean()) %>%        # get the (survey weighted) proportion for the numerator
  left_join(youth_bs %>%                                        # left join in the denominators
              group_by(geoid) %>%                                     # group by geo
              summarise(pop = survey_total(na.rm=T))) %>%              # get the weighted total for overall geo
  mutate(rate=rate*100,
         rate_moe = rate_se*1.645*100,    # calculate the margin of error for the rate based on se
         rate_cv = ((rate_moe/1.645)/rate) * 100, # calculate cv for rate
         count_moe = num_se*1.645, # calculate moe for numerator count based on se
         count_cv = ((count_moe/1.645)/num) * 100)  # calculate cv for numerator count


# rename race name columns as subgroup
total$subgroup = "Total"
total <- total %>% select(geoid, subgroup, everything())

aian <- aian %>% rename(subgroup = aian)
lat <- lat %>% rename(subgroup = latino)
nhpi <- nhpi %>% rename(subgroup = nhpi)
race <- race %>% rename(subgroup = race)
swana <- swana %>% rename(subgroup = swana)

# merge tables
d_long <- rbind(total, aian, lat, race, nhpi, swana) %>%
  filter(indicator == "under 5" & 
           subgroup != "Not AIAN" &
           subgroup != "Not Latinx" &
           subgroup != "Not NHPI" &
           subgroup != "Not SWANA" &
           subgroup != "AIAN placeholder" &
           subgroup != "NHPI placeholder" &
           subgroup != "Latinx placeholder")
           
d_long <- as.data.frame(d_long)
d_long <- d_long %>% select(geoid, everything()) %>% filter(is.na(geoid) == FALSE)

```

```{r}

#chart - THIS PART DOESN"T WORK YET - WAITING TO CHAT WITH JZ

#format for chart
d_chart <- d_long %>% filter(subgroup != "Total") %>%
  arrange(desc(num)) %>%
  mutate(across('subgroup', str_replace, ' NL', '')) %>%
  mutate(across('subgroup', str_replace, ' Alone or in Combination', '')) %>%
  rename(population = num) %>%
  mutate(group = "Children under 5")

# set options so I get commas for decimals
hcoptslang <- getOption("highcharter.lang")
hcoptslang$thousandsSep <- ","
options(highcharter.lang = hcoptslang)

hc <- d_chart %>%
  hchart('bar', hcaes(x = subgroup, y = population, group = group)) %>%
  hc_yAxis(title = list(text = "")) %>%
  hc_xAxis(title = list(text = "")) %>%
  hc_title(
    text = "LA County is home to more than 300,000 Latinx children under 5"
  ) %>%
   hc_exporting(
    enabled = TRUE, sourceWidth=900, sourceHeight=600,
    chartOptions=list(plotOptions=list(series=list(dataLabels=list(enabled=TRUE, format='{point.y:,.0f}')))),
    filename = "plot"
  ) %>%
  hc_subtitle(text = "Total number of children under age 5 by race") %>%
  hc_caption(text = "Source: ACS 2018-2022 5-year estimates. <br> AIAN = American Indian or Alaska Native. NHPI = Native Hawaiian or Pacific Islander. SWANA = Southwest Asian or North African. In this chart White, Asian, Black, Other and Two or More are non-Hispanic subgroups; AIAN, NHPI, SWANA, and Two or More are alone or in combination with other races. The sum of subgroups in this chart will exceed the total under 5 population.") %>%
  hc_legend(enabled = F)

hc
```

We include SWANA, similar to MENA (Middle Eastern and North African), because of recommendations that people in this group do not identify as non-Hispanic White (or other overlapping racial-ethnic groups) and experience different patterns of racial, ethnic, and religious discrimination than other groups.

## Why it matters?
Conversations about the under 5 population need to consider the needs of Latinx families, including:

 * dual language access for children, and materials for parents in both English and Spanish.

## Learn more
The importance of race and equity in the caring and educating of children under 5 cannot be overstated. See more in the [Data Narrative](https://apca-dev.dreamhosters.com/) for details.


