---
title: "Children under 5"
description: "The population of children under 5 decreased by 24% over the last 10 years." 
author:
  - name: Catalyst California
    url: catalystcalifornia.org
date: 2024-02-15
categories:
  - Indicators
output:
  distill::distill_article:
    css: css_theme.css 
    self_contained: false
preview: images/child0to5.PNG
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidycensus)
library(tidyverse)
library(scales)
library(highcharter)
library(magrittr)

```

```{r}
#make line chart of LA County under 5 population over time

#set variables
#v21 <- load_variables(2021, "acs1")
#View(v21)
pop_vars <- c(
  "B01001_003",
  "B01001_027")

#set years - no 2020 b/c of pandemic
years <- c(2013:2019, 2021:2022)
names(years) <- years


#get data
la_under5_by_year <- map_dfr(years, ~{
  get_acs(
  geography = "county",
  variables = pop_vars,
  state = "CA",
  county = "Los Angeles",
  survey = "acs1",
  year = .x
)
}, .id = "year")


#add male and female to get totals
la_under5_by_year <- la_under5_by_year %>% group_by(NAME, year) %>%
  summarize(population = sum(estimate),
            moe = moe_sum(moe, estimate = estimate, na.rm = FALSE))


#make 2020 value as average of 2019 and 2021 values
under5_2020 <- (la_under5_by_year[which(la_under5_by_year$year == 2019),]$population +
  la_under5_by_year[which(la_under5_by_year$year == 2021),]$population) / 2


#and add it to the rest
la_under5_by_year[nrow(la_under5_by_year) + 1,] = list("Los Angeles County, California", "2020", under5_2020, 0)
la_under5_by_year <- la_under5_by_year %>% arrange(year)
#View(la_under5_by_year)

```

```{r}

source("W:\\Project\\RDA Team\\First5LA\\First 5 LA Data Hub Year 5 2024\\R\\chart_prep\\f5la_chart_functions.R")


#TITLE 
title_text <- paste0("LA County households report 150,752 fewer children under 5 in 2022 than in 2013")

#SUBTITLE
subtitle_text <- paste0("Total number of children under age 5 by year")

#CAPTION
caption_text<- "Source: American Community Survey 1-year estimates table B01001 (2013-2022). Since 2020 estimates were not published, the 2020 estimate is the average of 2019 and 2021 estimates."

#TOOLTIP
#modify tooltips as needed 

tooltip_text <- paste0("<b>{point.population:,.0f}</b> children age 0-5 lived in Los Angeles County in <b>{point.year}</b>") #can always customize or add additional text to this as needed


#STEP 4 RUN the hchart FUNCTION 

graph <- fx_line(
    df = la_under5_by_year,
    x = 'year',
    y = 'population',
    chart_title = title_text,
    chart_subtitle = subtitle_text,
    chart_tooltip = tooltip_text,
    chart_caption = caption_text
)

graph

```

## Why it matters?
There's a disconnect between our values and the ability to raise a family in LA. We want to welcome children, but a nine-year decline in the under-five-population means families are choosing other options. Additionally,

 * fewer children means fewer State and Federal funds for schools and other critical infrastructure.

## Learn more
The decrease in children results from a number of factors that surveys and local experts can explain. See the [Data Narrative](https://apca-dev.dreamhosters.com/) for details.

