---
title: "Region 1 Indicators"
description: "Data related to children age 0 to 5 that align with Best Start needs and the new First 5 LA strategic plan within Region 1. Region 1 contains the East LA, Metro LA, South El Monte//El Monte, and Southeast LA Best Start geographies."
author:
  - name: Catalyst California
    url: catalystcalifornia.org
date: 2024-03-29
categories:
  - Region 1
  - Indicators
output:
  distill::distill_article:
    toc: true
    toc_float: true
    css: css_theme.css
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(highcharter)
options(scipen = 999)

# create connection for rda database and get data
source("W:\\RDA Team\\R\\credentials_source.R")
f5la_con <- connect_to_db("f5la_v2")
race_under5 <- dbGetQuery(f5la_con, "select * from data.f5la_2024_bs_raceethnicity_pums")
poverty_under5 <- dbGetQuery(f5la_con, "select * from data.f5la_2024_bs_poverty_pums")
rent_burden_under5 <- dbGetQuery(f5la_con, "select * from data.f5la_2024_bs_rent_burden_pums")
dbDisconnect(f5la_con)

source("W:\\Project\\RDA Team\\First5LA\\First 5 LA Data Hub Year 5 2024\\R\\chart_prep\\f5la_chart_functions.R")

```

```{r race_chart_function}

# make race chart function
make_race_chart <- function(df, best_start){
  d_chart <- df %>% 
  filter(geoid == best_start & subgroup != "Total") %>%
   arrange(desc(num)) %>%
   mutate(across('subgroup', str_replace, ' NL', '')) %>%
   mutate(across('subgroup', str_replace, ' Alone or in Combination', '')) %>%
   mutate(across('subgroup', str_replace, 'Two or More', 'Multiracial')) %>%
   rename(population = num) %>%
   mutate(group = "Children under 5")
  
#Get total pop under 5
total_under5 <- df %>% 
  filter(geoid == best_start & subgroup == "Total") %>%
  pull(num)

#Get under 5 population of race with most under 5
num_race_under5 <- d_chart[which.max(d_chart$population),]$population

#Get under 5 name of race with most under 5
max_race_under5 <- d_chart[which.max(d_chart$population),]$subgroup

#Calculate race's pct under 5 of total
d_chart$pct = d_chart$population / total_under5 * 100

#TITLE 
title_text <- paste0(max_race_under5, " children under 5 make up the largest racial-ethnic group in ", best_start)

#SUBTITLE
subtitle_text <- paste0("Number of children under 5 by race-ethnicity")

#CAPTION
caption_text <- caption_text<- paste0(acsnote,"<br>",racenote)

#TOOLTIP
tooltip_text <- paste0("<b>{point.population:.,1f}</b> {point.subgroup} children under 5 or 
                       <b>{point.pct:.1f}%</b> of all children under 5 in <b>{point.geoid}</b>")

#STEP 4 RUN the hchart FUNCTION 
graph <- fx_single(
    df = d_chart,
    x = 'subgroup',
    y = 'population',
    chart_title = title_text,
    chart_subtitle = subtitle_text,
    chart_tooltip = tooltip_text,
    chart_caption = caption_text
)
  
return(graph)
}

```

```{r poverty_chart_function, results=FALSE}

# make poverty chart function
make_poverty_chart <- function(df, best_start){
 d_chart <- df %>% filter(geoid == best_start) %>%
   arrange(desc(rate)) %>%
   mutate(across('subgroup', str_replace, ' NL', '')) %>%
   mutate(across('subgroup', str_replace, ' Alone or in Combination', '')) %>%
   mutate(across('subgroup', str_replace, 'Two or More', 'Multiracial')) %>%
   rename(population = num) %>%
   mutate(group = "Percent of children under 5 in poverty") %>%
  filter(subgroup != "Total")

# get Total poverty rate
total_poverty_under5 <<- df %>%
  filter(geoid == best_start & subgroup == "Total") %>%
  pull(rate)

#Get under 5 poverty rate of race with highest poverty rate
rate_poverty_under5 <<- d_chart[which.max(d_chart$rate),]$rate

#Get under 5 name of race with most under 5
max_rate_poverty_under5 <<- d_chart[which.max(d_chart$rate),]$subgroup

#TITLE
title_text <- paste0(max_rate_poverty_under5, " children under 5 are most likely to be in poverty in ", best_start)

#SUBTITLE
subtitle_text <- paste0("Percent of children under 5 in poverty")

#CAPTION
caption_text<- paste0(acsnote,"<br>",racenote) #can always customize or add additional text to this as needed

#TOOLTIP
tooltip_text <- paste0("<b>{point.rate:.1f}%</b> of <b>{point.subgroup}</b> children under 5 are in poverty or <b>{point.population:,.0f}</b> children under 5 in <b>{point.geoid}</b>")  

#STEP 4 RUN the hchart FUNCTION
graph <- fx_single_tot(
    df = d_chart,
    x = 'subgroup',
    y = 'rate',
    chart_title = title_text,
    chart_subtitle = subtitle_text,
    chart_tooltip = tooltip_text,
    chart_caption = caption_text,
    line_label = paste0(" Total ", round(total_poverty_under5,1),"%"),
    line_value = round(total_poverty_under5,1)
)
 
if(nrow(d_chart) > 1) { 
  return(graph) 
} else {
  cat(paste0("The ", best_start, " under 5 poverty rate is ", round(total_poverty_under5,1), "%."))
  cat(paste0("The only racial/ethnic group with statistically stable poverty rates is ",
               d_chart[which.max(d_chart$rate),]$subgroup, " whose poverty rate is ",
               round(rate_poverty_under5,1), "%."))  
}
      
#The `r best_start` under 5 poverty rate is `r round(total_poverty_under5,1)`%. The only racial/ethnic group with statistically stable poverty rates (from ACS microdata) is `r max_rate_poverty_under5`, whose poverty rate is `r round(rate_poverty_under5,1)`%.

}

```


```{r rent_burden_chart_function}

# make rent burden chart function
make_rent_burden_chart <- function(df, best_start){
 d_chart <- df %>% filter(geoid == best_start) %>%
   arrange(desc(rate)) %>%
   mutate(across('subgroup', str_replace, ' NL', '')) %>%
   mutate(across('subgroup', str_replace, ' Alone or in Combination', '')) %>%
   mutate(across('subgroup', str_replace, 'Two or More', 'Multiracial')) %>%
   rename(population = num) %>%
   mutate(group = "Percent of children under 5 in poverty") %>%
  filter(subgroup != "Total")

# get Total rent burden rate
total_rent_burden_under5 <- df %>% 
  filter(geoid == best_start & subgroup == "Total") %>%
  pull(rate)

#Get under 5 rent burden rate of race with highest rent burden rate
rate_rent_burden_under5 <- d_chart[which.max(d_chart$rate),]$rate

#Get under 5 name of race with highest rent burden rate
max_rate_rent_burden_under5 <- d_chart[which.max(d_chart$rate),]$subgroup

#TITLE 
title_text <- paste0(max_rate_rent_burden_under5, " children under 5 are most likely to be rent-burdened in ", best_start)

#SUBTITLE
subtitle_text <- paste0("Percent of children under 5 in rent-burdened households")

#CAPTION
caption_text<- paste0(acsnote,"<br>",racenote) #can always customize or add additional text to this as needed

#TOOLTIP
tooltip_text <- paste0("<b>{point.rate:.1f}%</b> of <b>{point.subgroup}</b> children under 5 are rent-burdened or  <b>{point.population:,.0f}</b> children under 5 in <b>{point.geoid}</b> ")


#STEP 4 RUN the hchart FUNCTION 
graph <- fx_single_tot(
    df = d_chart,
    x = 'subgroup',
    y = 'rate',
    chart_title = title_text,
    chart_subtitle = subtitle_text,
    chart_tooltip = tooltip_text,
    chart_caption = caption_text,
    line_label = paste0(" Total ", round(total_rent_burden_under5,1),"%"),
    line_value = round(total_rent_burden_under5,1)
)

return(graph)
}

```

# East LA

## Race-Ethnicity

```{r race1}

best_start <- "East LA"
make_race_chart(race_under5, best_start)

```


## Poverty

```{r poverty1, out.width="100%", out.height="100%", results = 'asis'}

best_start <- "East LA"
make_poverty_chart(poverty_under5, best_start)

```


## Rent Burden

```{r rb1}

best_start <- "East LA"
make_rent_burden_chart(rent_burden_under5, best_start)

```



# Metro LA

## Race-Ethnicity

```{r race2}

best_start <- "Metro LA"
make_race_chart(race_under5, best_start)

```

## Poverty

```{r poverty2, out.width="100%", out.height="100%", results = 'asis'}

best_start <- "Metro LA"
make_poverty_chart(poverty_under5, best_start)

```

## Rent Burden

```{r rb2 chart}

best_start <- "Metro LA"
make_rent_burden_chart(rent_burden_under5, best_start)

```


# South El Monte/El Monte

## Race-Ethnicity

```{r race3}

best_start <- "South El Monte/El Monte"
make_race_chart(race_under5, best_start)

```

## Poverty

```{r poverty3, out.width="100%", out.height="100%", results = 'asis'}

best_start <- "South El Monte/El Monte"
make_poverty_chart(poverty_under5, best_start)

```

## Rent Burden

```{r rb3}

best_start <- "South El Monte/El Monte"
make_rent_burden_chart(rent_burden_under5, best_start)

```


# Southeast LA

## Race-Ethnicity

```{r race4}

best_start <- "Southeast LA"
make_race_chart(race_under5, best_start)

```

## Poverty

```{r poverty4, out.width="100%", out.height="100%", results = 'asis'}

best_start <- "Southeast LA"
make_poverty_chart(poverty_under5, best_start)

```

## Rent Burden

```{r rb4 }

best_start <- "Southeast LA"
make_rent_burden_chart(rent_burden_under5, best_start)

```

# Why it matters
Understanding the racial-ethnic breakdown of an area is important to understanding its cultures and needs. Poverty and rent burden are indicators of economic need First 5 LA is interested in as it works to ensure all children prenatal to age 5 and their families have their basic needs met.

# Learn more
Learn more by reading the [First 5 LA 2024-2029 Strategic Plan](/https://www.first5la.org/wp-content/uploads/2023/11/First-5-LA-Strategic-Plan-2024-2029.pdf).

