---
title: "Licensed Child Care Centers and Large Family Child Care Homes with Under 5 Population"
description: |
  Numbers of licensed child care centers, licensed large family child care homes, capacity for each, and children under age 5 (with margin of error) by ZIP code. Zip codes used in analysis were defined by First 5 LA. Only facilities licensed by the California Department of Social Services are included.
author:
  - name: Catalyst California
    url: catalystcalifornia.org
date: 2025-01-22
categories:
  - Analysis
output:
  distill::distill_article:
    self_contained: false
preview: images/fire.jpg
---

This analysis was created by request of First 5 LA during the 2025 wildfires in Los Angeles in January 2025. The analysis was for specific zip codes requested by First 5 LA as part of their rapid response efforts. 

```{r, include=FALSE}

knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)

#### Load libraries ####

library(dplyr)
library(sf)
library(leaflet)
library(lubridate)
library(foreign)
library(tidycensus)
library(data.table)
library(janitor)

#### Add database connection ####

source("W:\\RDA Team\\R\\credentials_source.R")

con <- connect_to_db("la_fires")
con_shared <- connect_to_db("rda_shared_data")

options(scipen = 999)

# Load F5LA specific zip codes------------------

f5la_zips<-c("90272", "90265", "90290", "91101", "91103", "91104", "91105", "91106", "91107", "91001", "91024", "91011", "91342", "91344")

f5la_zips_df<-as.data.frame(f5la_zips)

```

```{r, include=FALSE}

# ANALYSIS: Licensed Child Care Centers ---------------------------------

# Read in data from postgres

#childcare<-st_read(con_shared, query="SELECT * FROM education.ccld_childcare_centers_2023", geom="geom")

# read in and format child care center data

childcare <- read.csv("W:/Data/Education/CCLD/CCLD 2024/ChildCareCenters01192025.csv", header = FALSE, sep = ",")
childcare <- childcare[,1:14]
names(childcare) <- childcare[1,]
childcare <- childcare[-1,]
childcare <- clean_names(childcare)
childcare$facility_capacity <- as.numeric(childcare$facility_capacity)

# Take child care df and filter only for F5LA zips, then calculate total numbers

childcare_f5la<-childcare%>%
  filter(facility_zip %in% f5la_zips & facility_status=="LICENSED")%>%
  as.data.frame()%>%
  group_by(facility_zip)%>%
  summarise(total_centers=n(),
            capacity_centers=sum(facility_capacity))


# ANALYSIS: Family Child Care Homes ------------------------------------


# Read in data from postgres

#familyhome<-dbGetQuery(con_shared, "SELECT * FROM education.ccld_family_childcare_homes_2023")

familyhome <- read.csv("W:/Data/Education/CCLD/CCLD 2024/CHILDCAREHOMEmorethan801192025.csv", header = FALSE, sep = ",")
familyhome <- familyhome[,1:14]
names(familyhome) <- familyhome[1,]
familyhome <- familyhome[-1,]
familyhome <- clean_names(familyhome)
familyhome$facility_capacity <- as.numeric(familyhome$facility_capacity)


# Take df and filter only for F5LA zips, then calculate total numbers

familyhome_f5la<-familyhome%>%
  filter(facility_zip %in% f5la_zips & facility_status=="LICENSED")%>%
  group_by(facility_zip)%>%
  summarise(total_family_homes=n(),
            capacity_family_homes=sum(facility_capacity))

# final table: join back f5la zips so we can see which zips had 0

familyhome_f5la<-f5la_zips_df%>%
  left_join(familyhome_f5la, by=c("f5la_zips"="facility_zip"))


# ANALYSIS: Age 0-5-------------------------------------------------

# age_f5la<-dbGetQuery(con_shared, "SELECT * FROM demographics.acs_5yr_s0101_multigeo_2022")%>%
#   filter(geolevel=="zcta" & name %in% f5la_zips)%>%
#   select(name, s0101_c01_002e, s0101_c01_002m	) # select columns of interest: zip, total pop under age 5, and MOE for total pop under age 5
# 
# # Analyze data
# 
# age_f5la<-age_f5la%>%
#   rename("zcta"="name")%>%
#   group_by(zcta)%>%
#   summarise(total_0_5=sum(s0101_c01_002e),
#             moe_0_5=moe_sum(s0101_c01_002m))

# identify ACS age under 5 variables to import

acs_variable <- load_variables(2023, "acs5/subject") %>% 
  filter(concept == "Age and Sex" & label == "Estimate!!Total!!Total population!!AGE!!Under 5 years")

tidy_variables <- c(paste0(as.character(acs_variable[1,1]), "E"), paste0(as.character(acs_variable[1,1]), "M"))

# import ACS data

age_0to5 <- get_acs(geography = "zcta",
                    variables = tidy_variables,
                    year = 2023,
                    dataset = "acs5/subject",
                    cache_table = TRUE)

# filter by F5LA ZIP codes and format for join

age_f5la <- subset(age_0to5, GEOID %in% f5la_zips) %>% 
  rename(zcta = GEOID,
         total_0_5 = estimate,
         moe_0_5 = moe) %>%
  select(-NAME, -variable)


# Final combine into single table ----------------------------------------

df<-childcare_f5la%>%
  rename("f5la_zips"="facility_zip")%>%
  left_join(familyhome_f5la)%>%
  left_join(age_f5la, by=c("f5la_zips"="zcta"))%>%
   mutate(across(everything(), ~replace(., is.na(.), 0))) # Replace all NA values with "N/A" for  visualizing


# Disconnect 

dbDisconnect(con)
dbDisconnect(con_shared)

``` 


```{r}

# Visualize final table

library(flextable)
library(officer)

ft<-flextable(df)%>%
  set_header_labels(f5la_zips = "Zip Code", total_centers = "Child Care Centers", capacity_centers = "Child Care Center Seats", total_family_homes="Large Family Child Care Homes", capacity_family_homes="Large Family Child Care Home Seats", total_0_5="Children Under Age 5", moe_0_5 = "MOE Under Age 5")%>%
  align(align = "center", part = "all") %>%
  autofit()

# Add borders

ft <- border_remove(x = ft)
ft <- hline(ft, part = "all", border =  fp_border(color = "gray"))

# Bold headers

ft <- bold(ft, part = "header")

ft

```

Data Sources:

*Child Care Centers and Large Family Child Care Homes* <br>
California Department of Social Services (CDSS) 2024. Data retrieved from https://www.ccld.dss.ca.gov/carefacilitysearch/DownloadData on 1/21/2024


*Children Under Age 5* <br>
American Community Survey (ACS) 5-year estimates 2019-2023, table S0101

