---
title: "Infant Deaths"
description: "Analysis of the infant death rate in Los Angeles county as part of the set of key data indicators related to children under age 5 that Catalyst California provides to First 5 LA."
author:
  - name: Catalyst California
    url: catalystcalifornia.org
date: 2024-03-06
categories:
  - Indicators
output:
  distill::distill_article:
    css: css_theme.css 
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidycensus)
library(tidyverse)
library(scales)
library(highcharter)
library(magrittr)
library(data.table)

```


```{r}
#make a line chart of infant deaths over liver births per 1,000 over time (infant mortality rate)

#get CHHS deaths. There are provisional deaths for 23-24, but the numbers are half 2022 or less so suspicious
deaths14_22 <- fread("https://data.chhs.ca.gov/datastore/dump/579cc04a-52d6-4c4c-b2df-ad901c9049b7?bom=True",
                        encoding="UTF-8")

deaths <- deaths14_22 %>% filter(County == "Los Angeles" & Strata_Name == "Under 1 year") %>% 
  group_by(Year) %>% summarize(Year = min(Year), Deaths = sum(Count))

#get CHHS births, comes in two sets of years
births60_22 <- fread("https://data.chhs.ca.gov/datastore/dump/d6c30e46-8618-407a-ba5a-bae308f86a1c?bom=True", 
                        encoding="UTF-8")

births <- births60_22 %>% filter(County == "Los Angeles" & Year >= 2014) %>% 
  group_by(Year) %>% summarize(Year = min(Year), Births = sum(Count, na.rm = TRUE))

mortality <- left_join(deaths, births, by = "Year") %>% mutate(mortality = Deaths/Births*1000)

```

```{r}

source("W:\\Project\\RDA Team\\First5LA\\First 5 LA Data Hub Year 5 2024\\R\\chart_prep\\f5la_chart_functions.R")


#TITLE 
title_text <- paste0("LA County Infant Mortality is Down from 2017")

#SUBTITLE
subtitle_text <- paste0("Number of infant deaths per 1,000 births by year")

#CAPTION
caption_text<- "Source: California Department of Public Health (2014-2022)"

#TOOLTIP
#modify tooltips as needed 

tooltip_text <- paste0("<b>{point.Deaths:,.0f}</b> infants under age 1 died in Los Angeles County in <b>{point.Year},</b><br> or <b>{point.mortality:.1f}</b> deaths per 1,000 live births.") #can always customize or add additional text to this as needed


#STEP 4 RUN the hchart FUNCTION 

graph <- fx_line(
    df = mortality,
    x = 'Year',
    y = 'mortality',
    chart_title = title_text,
    chart_subtitle = subtitle_text,
    chart_tooltip = tooltip_text,
    chart_caption = caption_text
)

graph

```


## Why it matters
While any infant death is one too many, a decline in numbers and rates of infant deaths is something to celebrate. Fewer infant deaths means more healthy infants and families.

## Learn more
Areas of concern remain for infants. Life/death is unfortunately not the only measure of child welfare. Antelope Valley partners highlight concerns over child welfare. See their [State of the Child report](https://www.all4kids.org/wp-content/uploads/2023/08/Region5_StateOfTheChildReport_2023_Final_Web2-1.pdf) for details.
