---
title: "Real Cost Measure"
description: |
  Updated analysis of Real Cost Measure by race in Los Angeles county (updated in 2023), as part of Catalyst California's annual update of key indicators related to families with children age 0-5. 
author:
  - name: Catalyst California
    url: catalystcalifornia.org
date: 2024-03-20
categories:
  - Indicators
output:
  distill::distill_article:
    self_contained: false
preview: images/RCM.png
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(data.table)
library(highcharter)
library(stringr)
library(RPostgreSQL)
library(reshape2) 
library(scales)

source("W:\\RDA Team\\R\\credentials_source.R")
con <- connect_to_db("rda_shared_data")

source("W:\\Project\\RDA Team\\First5LA\\First 5 LA Data Hub Year 5 2024\\R\\chart_prep\\f5la_chart_functions.R")
```

```{r, include=FALSE}

# read in 2023 RCM data

rcm<-dbGetQuery(con, "SELECT * FROM economic.uw_2023_county_state_real_cost_measure")

# filter for columns and geos of interest

df<-rcm%>%
  filter(county == 'Los Angeles')%>%
  select(14:25)

# pivot table longer
 df<-df%>%
   pivot_longer(1:12, names_to = "variable", values_to = "value")

#make data long for both pct and number

pct <- df%>%
  filter(grepl("pct", variable))%>%
  rename("pct_var"="variable",
         "pct_value"="value")%>%
mutate(pct_value=pct_value*100)

num<-df%>%
  filter(grepl("num", variable))%>%
  rename("num_var"="variable",
         "num_value"="value")

#combine

rcm_race<-cbind(num, pct)

#insert value for pct hh aian from dashboard for 2023: https://public.tableau.com/app/profile/hgascon/viz/TheRealCostMeasureinCalifornia2023/RealCostDashboard?publish=yes

rcm_race<-rcm_race%>%
  mutate(pct_value=ifelse(pct_var %in% "pct_aian_hh_below_rcm", 50, pct_value))

#now remove rows with the NA values because we can't graph

rcm_race<-rcm_race[-c(6), ] 

#now create new variable column with nicer labeling for the graph

rcm_race<-rcm_race%>%
  mutate(variable=ifelse(pct_var %in% 'pct_white_hh_below_rcm', 'White',
                         ifelse(pct_var %in% 'pct_black_hh_below_rcm', 'Black',
                                ifelse(pct_var %in% 'pct_asian_hh_below_rcm', 'Asian',
                                                ifelse(pct_var %in% 'pct_latino_hh_below_rcm', 'Latinx', ifelse(pct_var %in% 'pct_other_hh_below_rcm', 'Other', "American Indian/Alaskan Native"))))))%>%
  mutate(across(is.numeric, ~ round(., 1)))%>%
  arrange(-pct_value)

```

```{r}

# graph

#TITLE 
title_text <- paste0("Latinx and AIAN households are 2x more likely to earn less than the Real Cost Measure compared to White households")

#SUBTITLE
subtitle_text <- paste0("Percent of households with young children earning below the Real Cost Measure in Los Angeles county")

#CAPTION
caption_text<- paste0("The Real Cost Measure in California 2023. United Ways of California. June 2023. https://www.unitedwaysca.org/realcost.") #can always customize or add additional text to this as needed

#TOOLTIP
#modify tooltips as needed 

tooltip_text <- paste0("<b>{point.pct_value:.1f}%</b> of <b>{point.variable}</b> households with young children earn below the Real Cost Measure, or a total of <b>{point.num_value:,.0f}</b> households")


#STEP 4 RUN the hchart FUNCTION 

graph <- fx_single(
    df = rcm_race,
    x = 'variable',
    y = 'pct_value',
    chart_title = title_text,
    chart_subtitle = subtitle_text,
    chart_tooltip = tooltip_text,
    chart_caption = caption_text
)

graph


```

The data below visualizes the Real Cost Measure budget for a family in Los Angeles county with two children, one child age 0-5 and one school-aged child. This is visualized alongside the Los Angeles county median household income. All data is from 2021. 

```{r, fig.width=10,fig.height=9}

# note this still has not been updated and is from 2021. I took this code from: "W:\Project\RDA Team\First5LA\V2.0 Report\R\Real Cost Measure 2023.rmd"

#####Create data frame for RCM budget for LA County family of 4 (two adults, one preschooler, one school child)

rcm_la <-data.frame("measure"=c("Real Cost Measure \nAnnual Household Budget", 
                                "Real Cost Measure \nAnnual Household Budget", 
                                "Real Cost Measure \nAnnual Household Budget", 
                                "Real Cost Measure \nAnnual Household Budget", 
                                "Real Cost Measure \nAnnual Household Budget", 
                                "Real Cost Measure \nAnnual Household Budget", 
                                "Real Cost Measure \nAnnual Household Budget", 
                               "Median Household Income"),
                    expense=c("Housing", 
                              "Child Care",
                              "Food", 
                              "Health Care", 
                              "Transportation",
                              "Miscellaneous", 
                              "Taxes", 
                              "Median Household \nIncome"),
                    "amount"=c(24696, 
                               18708, 
                               12096, 
                               10080, 
                               10884, 
                               7644, 
                               14520, 
                               76367))


#customize the order for the stacked bar graph
rcm_la$expense <- factor(rcm_la$expense,levels = c("Child Care", "Food", "Health Care", "Housing", "Transportation", "Taxes", "Miscellaneous", "Median Household \nIncome"))

#####Graph

plot<-ggplot(rcm_la, aes(fill=expense, y=amount, x=measure, label=amount))+
  
  geom_bar(position="stack", stat="identity")+
  
  geom_text(aes(label=paste0("$", format(amount, big.mark=","))), size = 6, position = position_stack(vjust = 0.5))+
   
  scale_y_continuous(name="", labels=dollar_format(), breaks=seq(0,95000,10000))+
  
  scale_x_discrete(name="")+
  
  labs(fill="",
       caption=paste0("\n",str_wrap("Data Sources: Real Cost Measure annual budget for a family with two adults, one preschooler, and one school-aged child living in Los Angeles County from Struggling to Stay Afloat: The Real Cost Measure in California 2021, United Ways of California. Median household income from American Community Survey 2017-2021 5-year estimates table B19013.", 120)))+
  
   scale_fill_manual(values=c("#332985", "#009CDB", "#22BCB8", "#54B847","#F58326","#EC098C", "#EF4034", "grey"))+

  ggtitle("Real Cost Budget for Family with Children in\nLos Angeles County")+
  
 theme(panel.background = element_blank(),
        plot.title =  element_text(hjust = 0, size = 32, colour = black),
        plot.caption = element_text(hjust = 0, size = 13, colour = black),
        legend.text=element_text(size=15),
       axis.text.x = element_text(size = 18),
        axis.text.y = element_text(size = 14),
        axis.ticks = element_blank(),
        # grid line style
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(size = 0.25),
        panel.border = element_blank(),
        panel.grid.major.y = element_blank(),
        axis.line=element_blank()) 


plot

 ggsave(plot=plot, filename="RCM.png",
        path=paste0("W:\\Project\\RDA Team\\First5LA\\First 5 LA Data Hub Year 5 2024\\Github\\JZ\\f5la_datahub\\f5la_datahub\\_posts\\2024-03-20-real-cost-measure\\images"),
        height=8.5, width=11.2, units="in", dpi=150)


```

## Why it matters
First 5 LA's first strategic plan goal is that children prenatal to age 5 and their families have their basic needs met, which includes economic security. Economic security is foundational to reducing the annual percent of households with prenatal to 5 living poverty and experiencing housing insecurity, which are also part of the first strategic plan goal. The Real Cost Measure is a more holistic indicator that measures economic security in the context of families with young children, and what it costs to meet the essential needs of young children. 

## Learn more
See data for other indicators of basic needs in the [First 5 LA 2024-2029 Strategic Plan](https://www.first5la.org/wp-content/uploads/2023/11/First-5-LA-Strategic-Plan-2024-2029.pdf) and see related economic indicators and indicators by race/ethnicity in the [Data Narrative](https://apca-dev.dreamhosters.com/).

To learn more about Guaranteed Basic Income (GBI) efforts in Los Angeles county and how GBI improves outcomes for children and families, see the Catalyst California GBI presentation from the First 5 LA GBI Learning Dialogue [here](https://catalystcalifornia.github.io/f5la_datahub/f5la_datahub/docs/posts/2024-01-23-gbi-learning-dialogue-presentation/)

```{r, include = FALSE}
#Close postgres connection
dbDisconnect(con) 
```