---
title: "Rent Burden of Households with Children under 5"
description: "Updated analysis of households with children age 0-5 experiencing rent burden in Los Angeles county (updated in 2024), as part of Catalyst California's annual update of indicators."
author:
  - name: Catalyst California
    url: catalystcalifornia.org
date: 2024-06-05
categories:
  - Indicators
output:
  distill::distill_article:
    self_contained: false
preview: images/rentburden.PNG
---


```{r setup, include=FALSE}

# get LA County rent-burdened households with children under5 by race
# Data Dictionary: https://www2.census.gov/programs-surveys/acs/tech_docs/pums/data_dict/PUMS_Data_Dictionary_2018-2022.pdf

library(tidyverse)
library(data.table)
library(readxl)
library(tidycensus)
library(srvyr)
library(highcharter)
library(stringr)

#SOURCE from the script that has: styling, packages, dbconnection, colors
source("W:\\RDA Team\\R\\credentials_source.R")

con <- connect_to_db("f5la_v2")

source("W:\\Project\\RDA Team\\First5LA\\First 5 LA Data Hub Year 5 2024\\R\\chart_prep\\f5la_chart_functions.R")


```

```{r data, include=FALSE}

 source("W:\\Project\\RDA Team\\First5LA\\Ad-hoc Research\\Laura ask\\Rent Burden\\rent_burden_by_race.R")

```

```{r, echo=FALSE,out.width="100%", out.height="100%"}

 #format for chart
 d_chart <- d_long %>% 
   arrange(desc(rate)) %>%
   mutate(across('subgroup', str_replace, ' NL', '')) %>%
   mutate(across('subgroup', str_replace, ' Alone or in Combination', '')) %>%
   rename(population = num) %>%
   mutate(group = "Percent of children under 5 in rent-burdened households") %>%
  filter(subgroup != "Total")

#TITLE 
title_text <- paste0("SWANA, Black, and Latinx households with children under 5 are more likely to be rent burdened")

#SUBTITLE
subtitle_text <- paste0("Percent of households with children under 5 experiencing rent burden in Los Angeles county")

#CAPTION
caption_text<- paste0(acsnote,"<br>",racenote) #can always customize or add additional text to this as needed

#TOOLTIP
#modify tooltips as needed 

tooltip_text <- paste0("<b>{point.rate:.1f}%</b> of <b>{point.subgroup}</b> households with children under 5 experience rent burden,<br> or a total of <b>{point.pop:,.0f}</b> households with children under 5")


#STEP 4 RUN the hchart FUNCTION 

#STEP 4 RUN the hchart FUNCTION 
total_rate <- round(d_long$rate[ which(d_long$subgroup=="Total")],1)

graph <- fx_single_tot(
    df = d_chart,
    x = 'subgroup',
    y = 'rate',
    chart_title = title_text,
    chart_subtitle = subtitle_text,
    chart_tooltip = tooltip_text,
    chart_caption = caption_text,
    line_label= paste0("Total: ", total_rate, "%"),
    line_value= total_rate
)

graph

```

## Why it matters
First 5 LA's first strategic plan goal is that children prenatal to age 5 and their families have their basic needs met, which includes secure housing for essential psychological safety. This goal has an objective to reduce the annual percent of households with children prenatal through age 5 who experience housing insecurity. Rent burden is not the sole cause of housing insecurity, but it is a key contributor.

## Learn more
See data for other indicators of basic needs in the [First 5 LA 2024-2029 Strategic Plan](https://www.first5la.org/wp-content/uploads/2023/11/First-5-LA-Strategic-Plan-2024-2029.pdf) and see related housing indicators and indicators by race/ethnicity in the [Data Narrative](https://apca-dev.dreamhosters.com/).

```{r, include = FALSE}
#Close postgres connection
dbDisconnect(con) 
```

