---
title: "Region 4 Indicators"
description: "Data related to children age 0 to 5 that align with Best Start needs and the new First 5 LA strategic plan within Region 4. Region 4 contains the Central Long Beach and Wilmington geographies."
author:
  - name: Catalyst California
    url: catalystcalifornia.org
date: 2024-03-29
categories:
  - Region 4
  - Indicators
output:
  distill::distill_article:
    toc: true
    toc_float: true
    css: css_theme.css
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(highcharter)
options(scipen = 999)

# create connection for rda database and get data
source("W:\\RDA Team\\R\\credentials_source.R")
f5la_con <- connect_to_db("f5la_v2")
race_under5 <- dbGetQuery(f5la_con, "select * from data.f5la_2024_bs_raceethnicity_pums")
poverty_under5 <- dbGetQuery(f5la_con, "select * from data.f5la_2024_bs_poverty_pums")
poverty_under5_dis <- dbGetQuery(f5la_con, "select * from f5la_2024_bs_poverty_asian_dis_pums")
rent_burden_under5 <- dbGetQuery(f5la_con, "select * from data.f5la_2024_bs_rent_burden_pums")
rent_burden_under5_dis <- dbGetQuery(f5la_con, "select * from data.f5la_2024_bs_rent_burden_asian_dis_pums")
dbDisconnect(f5la_con)

source("W:\\Project\\RDA Team\\First5LA\\First 5 LA Data Hub Year 5 2024\\R\\chart_prep\\f5la_chart_functions.R")

# 5.23.24: add source to the R script version of this that has all of the code to get it to run

source("W:\\Project\\RDA Team\\First5LA\\First 5 LA Data Hub Year 5 2024\\R\\analysis\\region_4_indicators.R")



```


# Central Long Beach 

## Race-Ethnicity

```{r race1}

best_start <- "Central Long Beach"
make_race_chart(race_under5, best_start)

```


## Poverty

```{r poverty1, out.width="100%", out.height="100%", results = 'asis'}

best_start <- "Central Long Beach"
make_poverty_chart(poverty_under5, best_start)

```

### Disaggregating the Asian Subgroup

Statistics of Asian populations can often mask stark differences between people of different ancestries. Central Long Beach partners ask that we disaggregate these data when possible because of the Asian populations in the area. Unfortunately, there are no stable estimates of children under 5 in poverty by disaggregated Asian ancestry.

```{r, echo=FALSE,out.width="100%", out.height="100%"}

 #format for chart
 d_chart_asian <- poverty_under5_dis %>% rename(subgroup = asian) %>%
   filter(indicator =="below poverty" & pop >= 100 & rate_cv <= 40 & 
            geoid == "Central Long Beach" & subgroup != "Not Asian") %>%
   arrange(desc(rate)) %>%
   rename(population = num) %>%
   mutate(group = "Percent of children under 5 in poverty")

#TITLE 
title_text <- paste0("Mongolian children under 5 are most likely to be in poverty among Asians")

#SUBTITLE
subtitle_text <- paste0("Percent of children under 5 in poverty in Los Angeles County")

#CAPTION
caption_text<- paste0(acsnote,"<br>",racenote) #can always customize or add additional text to this as needed

#TOOLTIP
#modify tooltips as needed 

tooltip_text <- paste0("<b>{point.rate:.1f}%</b> of <b>{point.subgroup}</b> children under 5 in poverty, or <b>{point.population:,.0f}</b> children under 5")


#STEP 4 RUN the hchart FUNCTION 

# NO STABLE DATA TO GRAPH
# graph <- fx_single_tot(
#     df = d_chart_asian,
#     x = 'subgroup',
#     y = 'rate',
#     chart_title = title_text,
#     chart_subtitle = subtitle_text,
#     chart_tooltip = tooltip_text,
#     chart_caption = caption_text,
#     line_label="Asian Total: 8.1%",
#     line_value="8.1"
# )

#graph

```




## Rent Burden

```{r rb1}

best_start <- "Central Long Beach"
make_rent_burden_chart(rent_burden_under5, best_start)

```

### Disaggregating the Asian Subgroup

Statistics of Asian populations can often mask stark differences between people of different ancestries. Central Long Beach partners ask that we disaggregate these data when possible because of the Asian populations in the area. See chart below for the Asian disaggregation of rent burden.

```{r, echo=FALSE,out.width="100%", out.height="100%"}

 #format for chart
 d_chart_asian <- rent_burden_under5_dis %>% rename(subgroup = asian) %>%
   filter(indicator =="burdened" & pop >= 100 & rate_cv <= 40 & 
            subgroup != "Not Asian" & geoid == "Central Long Beach") %>%
   arrange(desc(rate)) %>%
   rename(population = num) %>%
   mutate(group = "Percent of children under 5 in rent-burdened households")

#TITLE 
title_text <- paste0("Filipino and Cambodian children under 5 are most likely to be rent burdened among Central Long Beach Asians")

#SUBTITLE
subtitle_text <- paste0("Percent of children under 5 in rent-burdened households")

#CAPTION
caption_text<- paste0(acsnote,"<br>",racenote) #can always customize or add additional text to this as needed

#TOOLTIP
#modify tooltips as needed 

tooltip_text <- paste0("<b>{point.rate:.1f}%</b> of <b>{point.subgroup}</b> children under 5 experience rent burden, or a total of <b>{point.pop:,.0f}</b> children under 5")


#STEP 4 RUN the hchart FUNCTION 

graph <- fx_single_tot(
    df = d_chart_asian,
    x = 'subgroup',
    y = 'rate',
    chart_title = title_text,
    chart_subtitle = subtitle_text,
    chart_tooltip = tooltip_text,
    chart_caption = caption_text,
    line_label = " Asian Total: 45.9%",
    line_value = 45.9
)

graph

```




# Wilmington

## Race-Ethnicity

```{r race2}

best_start <- "Wilmington"
make_race_chart(race_under5, best_start)

```

## Poverty

```{r poverty2, out.width="100%", out.height="100%", results = 'asis'}

best_start <- "Wilmington"
make_poverty_chart(poverty_under5, best_start)

```

## Rent Burden

```{r rb2 chart}

best_start <- "Wilmington"
make_rent_burden_chart(rent_burden_under5, best_start)

```


# Why it matters
Understanding the racial-ethnic breakdown of an area is important to understanding its cultures and needs. Poverty and rent burden are indicators of economic need First 5 LA is interested in as it works to ensure all children prenatal to age 5 and their families have their basic needs met.

# Learn more
Learn more by reading the [First 5 LA 2024-2029 Strategic Plan](/https://www.first5la.org/wp-content/uploads/2023/11/First-5-LA-Strategic-Plan-2024-2029.pdf).

